### 2019-06-07 01_health_data.Rmd
Created health outcomes dataset

---
title: "Health Data"
author: "Greg Macfarlane"
date: "6/6/2019"
output: html_document
---

```{r setup}
library(leaflet)
library(tidyverse)
library(tidycensus)
library(RSocrata)
library(sf)
options(tigris_use_cache=TRUE)
```

In this document we collect the health and socioeconomic data necessary for the
project from the American Community Survey and the Centers for Disease Control.

## American Community Survey
We will use the `tidycensus` package from Kyle Walker to get the datatables in 
our analysis region, which is the five boroughs of New York City.

```{r region}
# The study area includes the five boroughs of New York City
counties <- c("081", "047", "061", "005", "085")
names(counties) <- c("Queens", "Brooklyn", "Manhattan", "Bronx", "Staten Island")
```

We need to get lots of variables from the American Community Survey 5-year data.
This vector defines the variable names as well as the table/cell in the Census
API that each variable is drawn from. The comments help to identify which
table name and population type they come from.
```{r acsvariables}
variables <- c(
  "population" = "B25008_001", # TOTAL POPULATION IN OCCUPIED HOUSING UNITS BY TENURE
  "housing_units" = "B25001_001", # HOUSING UNITS
  "households" = "B19001_001", #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  # Employment status
  "men_fulltime" = "B23022_004",
  # Estimate!!Total!!Male!!Worked in the past 12 months!!Usually worked 35 or more hours per week
  "women_fulltime" = "B23022_028",
  #Estimate!!Total!!Female!!Worked in the past 12 months!!Usually worked 35 or more hours per week
  # LIVING ARRANGEMENTS OF ADULTS 18 YEARS AND OVER BY AGE
  # everyone who
  "adults" = "B09021_001",
  "lives_alone" = "B09021_002",
  "child_of_hhr" = "B09021_005",
  "other_rel" = "B09021_006",
  "other_nonrel" = "B09021_007",
  # EDUCATIONAL ATTAINMENT BY EMPLOYMENT STATUS FOR THE POPULATION 25 TO 64 YEARS
  "college" = "B23006_016",
  # RACE
  "black" = "B02001_003",
  "asian" = "B02001_005",
  "pacific" = "B02001_006",
  "nativeam" = "B02001_004",
  # HISPANIC OR LATINO ORIGIN BY SPECIFIC ORIGIN
  # The number of hispanic individuals needs to be drawn from a different table.
  "hispanic" = "B03001_003",
  #MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "income" = "B19013_001",
  #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "inc_0010" = "B19001_002",  "inc_1015" = "B19001_003", "inc_1520" = "B19001_004",
  "inc_2025" = "B19001_005", "inc_2530" = "B19001_006", "inc_3035" = "B19001_007",
  "inc_125"  = "B19001_015", "inc_150"  = "B19001_016", "inc_200"  = "B19001_017",
  #AGE BY RATIO OF INCOME TO POVERTY LEVEL IN THE PAST 12 MONTHS.
  # we really only need the numbers of individuals in each age group, and the groups
  # are more coarse here than in the sex by age table.
  "age_0005" = "B17024_002", "age_0611" = "B17024_015", "age_1217" = "B17024_028",
  "age_1824" = "B17024_041", "age_2534" = "B17024_054",
  "age_3544" = "B17024_067", "age_4554" = "B17024_080", "age_5564" = "B17024_093",
  "age_6574" = "B17024_106",
  "age_75up" = "B17024_119"
)
```

Now that we've defined the necessary variables and geographic scope, we can pull
the data from the API. Because the data coming from census are counts, we need 
to transform the numbers into percents of the relevant population denominator.

```{r acs}
acs <- get_acs(geography = "tract", variables = variables, year = 2017,
               state = "NY", county = counties, geometry = TRUE) %>%
  select(-moe) %>%
  spread(variable, estimate) %>%
  filter(adults > 0) %>%
  mutate(area = as.numeric(st_area(geometry) * 1e-6)) %>%
  # area is in m^2, change to km^2
  transmute(
    geoid = GEOID,
    county_fips = substr(GEOID, 3, 5),
    borough = case_when(
      county_fips == "081" ~ "Queens",
      county_fips == "047" ~ "Brooklyn",
      county_fips == "061" ~ "Manhattan",
      county_fips == "005" ~ "Bronx",
      county_fips == "085" ~ "Staten Island",
      TRUE ~ as.character(NA)
    ),
    population, households, housing_units, adults,
    density = households / area,
    income,
    # many of the variables come in raw counts, but we want to consider
    # them as shares of a relevant denominator.
    lowincome    = 100 * (inc_0010 + inc_1015 + inc_1520 + inc_2530 +
                            inc_3035) / households,
    highincome   = 100 * (inc_125 + inc_150 + inc_200) / households,
    fulltime     = 100 * (men_fulltime + women_fulltime) / adults,
    college      = 100 * college / adults,
    single       = 100 * (lives_alone + child_of_hhr + other_rel +
                            other_nonrel) / adults,
    youth        = 100 * (age_0005 + age_0611 + age_1217) / population,
    young_adults = 100 * (age_1824 + age_2534) / population,
    seniors      = 100 * (age_6574 + age_75up) / population,
    black        = 100 * black / population,
    asian        = 100 * asian / population,
    hispanic     = 100 * hispanic / population,
    other        = 100 * (nativeam + pacific) / population
  )
```

## 500 Cities
The Centers for Disease Control and Prevention makes small-area estimates on
key health indicators available through its 500 Cities data program (CDC5002016).
The indicators are multilevel aggregations and imputations of BRFSS responses
(Wang2018, Wang2017), and have been recently used to study the tract-level
link between gentrification and urban health (Gibbons2018). We use two
indicators as our dependent variables in this study: the share of adults
in a Census tract who are obese, and the share of adults who participate in no
leisure-time physical activity. To improve clarity in our interpretation, we use
the complement of the second variable --- the share of tract adults who
participate in {\em some} physical activity --- even if the amount is not
sufficient to affect overall health. Both indicators are estimated for the year
2016.

The CDC 500 cities project makes its data available through the Socrata API.
The url endpoints for each data object are different, though we have identified
them as below. We pre-filter the data to New York to save ourselves and the
API a bit of time. For each URL endpoint, get the data through Socrata (you will
need your own Socrata API key), filter to the FIPS codes we specified above, and
trim to only the data values we want to keep.

```{r cdc500}
cdc_urls <- list(
  "physact" = "https://chronicdata.cdc.gov/resource/axqz-rkkp.json?stateabbr=NY",
  "obesity" = "https://chronicdata.cdc.gov/resource/bjvu-3y7d.json?stateabbr=NY"
)

cdc500 <- lapply(cdc_urls, function(url){
  read.socrata(url, app_token = Sys.getenv("CDC_KEY"),
               stringsAsFactors = FALSE) %>%
    tbl_df() %>%
    filter(tractfips %in% acs$geoid) %>%
    transmute( geoid = tractfips, value = as.numeric(data_value ), year = year)
}) %>%
  bind_rows(.id = "measure") %>%
  spread(measure, value) %>%
  mutate(physact = 100 - physact)
```

Save the dataset for further analysis.
```{r writeout}
tracts <- left_join(acs, cdc500, by = "geoid") %>%
  filter(!is.na(obesity), !is.na(income)) %>% 
  write_rds("data/tracts.rds")
```