### 2019-06-07 02_parks_data.Rmd
Recreate parks dataset
Instead of using the geojson Nico downloaded and manipulated manually, we are
working directly from NYC's open data API. I also made a change by adding
cemeteries and combining shapes that belong to a common park, and removed
tennis courts and baseball diamonds, etc. that are part of a larger facility.

---
title: "Parks Data"
author: "Greg Macfarlane"
date: "6/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(leaflet)
library(tidyverse)
library(RSocrata)
library(leaflet)
library(sf)
```

In this document we get the parks data from the New York City open data
repository and do some cleaning. First, we get the parks dataset. The original
data dictionary is available at 
[on Github](https://github.com/CityOfNewYork/nyc-planimetrics/blob/master/Capture_Rules.md#open-space)
```{r parks}
parks_url <- "https://data.cityofnewyork.us/api/geospatial/y6ja-fw4f?method=export&format=GeoJSON"
parks <- st_read(parks_url, stringsAsFactors = FALSE) %>%
  mutate(shape_area = as.numeric(shape_area))
```

## Cleaning
The map of Flushing Meadows in Queens below shows some of the difficulties of this dataset:
  - First, the park is constructed of multiple outlines that limits the size of 
  the total park. This makes Flushing Meadows appear smaller than Central Park, 
  which is a single large park. 
  - Second, the dataset includes tennis parks and baseball fields as separate 
  facilities.
  - Third, parking lots for stadia (in this case Citi Field) are included as 
  open park space. We should remove these as they are categorically unlike other
  green spaces.
  - Fourth, important green spaces like cemeteries are in a different dataset.

```{r fm1}
leaflet(st_buffer(parks %>% filter(grepl("Flushing Meadows", park_name)), dist = 0.000001)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(label = ~as.character(park_name), stroke = 0.5)
```

First, we manually remove the park facilities corresponding to stadia and their
associated parking lots, and then filter out facilities that are not explicitly
park boundaries, as identified by the code `4980`.

```{r notparks}
notparks <- c(9498000833, 14498000872, 14498000794, # citi field, Yankee stadium, garage
              9498000631)  # USTA Billie Jean Tennis Center
parks <- parks %>% 
  filter(feat_code == 4980) %>%
  filter(!(source_id %in% notparks))
```

```{r fm2}
leaflet(st_buffer(parks %>% filter(grepl("Flushing Meadows", park_name)), dist = 0.000001)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(label = ~as.character(park_name), stroke = 0.5)
```

With those areas removed, we build a 30-foot buffer around each park and 
dissolve interior boundaries. This will make parks that are effectively the same
facility a single polygon feature. We also re-calculate the area of the parks in
acres and remove parks smaller than half an acre.

```{r dissolve}
dissolved <- parks %>%
  #filter(grepl("Flushing Meadows", park_name)) %>%
  st_transform(3628) %>%
  st_buffer(30) %>%
  group_by(parknum) %>%
  summarise(area = sum(shape_area),
            park_name = park_name[1]) %>%
  mutate(sqft = st_area(.), acres = units::set_units(sqft, acres)) 
```

```{r fm3}
leaflet(st_transform(dissolved, 4326) ) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5)
```

### Cemeteries
Some important green spaces that can be used for physical activity but are not
parks include cemeteries. These are stored in a separate dataset called Open
Space (Other) with the feature code `2500`.

```{r cemeteries}
cemeteries_url <- "https://data.cityofnewyork.us/api/geospatial/pckb-8r2z?method=export&format=GeoJSON"
cemeteries <- st_read(cemeteries_url) %>%
  filter(feature_co == 2500)
```

```{r d_cemeteries}
dissolved_cemeteries <- cemeteries %>%
  st_transform(3628) %>%
  st_buffer(5) %>%
  mutate(sqft = st_area(.), acres = units::set_units(sqft, acres))
```

```{r cemeteries_map}
leaflet(st_transform(dissolved_cemeteries, 4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5)
```

Now we can join the cemeteries data to the parks data and write out the results
```{r write}
rbind(
  dissolved %>% transmute(name = park_name, acres, type = "park"),
  dissolved_cemeteries %>% transmute(name, acres, type = "cemetery")
) %>%
  filter(as.numeric(acres) > 0.5) %>%
  st_simplify(preserveTopology = TRUE) %>%
  write_rds("data/open_spaces.rds")
```