### 2019-06-17 03_logsums.Rmd
Calculate the distance and size-based logsums

---
title: "Logsums"
author: "Greg Macfarlane"
date: "6/7/2019"
output: html_document
---

```{r setup}
library(tidyverse)
library(sf)
library(spdep)
library(leaflet)
```

In this document we calculate the distance between each tract in New York City 
and each park to help calculate a logsum-based measure of accessibility. We 
cleaned and gathered the data in the previous two pages.

```{r load}
open_spaces <- read_rds("data/open_spaces.rds")
tracts <- read_rds("data/tracts.rds")
```

We are going to do a point-to-polygon distance calculation. The geometric centroid
of the tracts are not satisfying, so we will use the population-weighted centroid
instead.

```{r pop_weighted_centroids}
pop_weighted_centroids <- tracts %>% tbl_df() %>%
  dplyr::select(geoid) %>%
  left_join(
    read_csv("https://www2.census.gov/geo/docs/reference/cenpop2010/tract/CenPop2010_Mean_TR36.txt") %>%
      transmute(geoid = str_c(STATEFP, COUNTYFP, TRACTCE), LATITUDE, LONGITUDE)
  ) %>%
  # convert to sf and project into the feet-based projection used in the
  # parks dataset.
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(3628)
```

The population-weighted centroid also serves as the orientation for the spatial
weights matrix.

```{r W}
# determine which tracts are neighbors of each other. Need to be less than
# 1.8 miles away from the centroids
tracts.dnn <- dnearneigh(as(pop_weighted_centroids, "Spatial"), 0, 1.8 * 5280)
# calculate the inverse distance to use as a weight on the neighbor connection
dists <- nbdists(tracts.dnn, as(pop_weighted_centroids, "Spatial"))
dists.inv <- lapply(dists, function(x) 1 / x)
# get the spatial weights matrix and its trace to use in any impact analyses
W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv,
              zero.policy = TRUE, style = "W")
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
print(W, zero.policy = TRUE)
```

## Distances
We want to make sure that the distance between a park and a tract centroid is
based on the edge of the polygon instead of the centroid thereof, so that spaces
at the corners of Central park, for instance, don't appear as longer than they should be.
It seems that things are working appropriately.

```{r test_measurement}
central_park <- open_spaces %>% filter(name == "Central Park")
pop_weighted_centroids$central_park <- log(as.numeric(st_distance(pop_weighted_centroids, central_park)))

dmeasure <- colorNumeric(palette = "Blues", domain = pop_weighted_centroids$central_park)
leaflet(pop_weighted_centroids %>% st_transform(4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addCircleMarkers(color = ~dmeasure(central_park))
```

Now we can calculate everything.
```{r distances}
# distances from all tract centroids to all parks
distances <- st_distance(pop_weighted_centroids, open_spaces, byid = TRUE) %>%
  units::set_units(miles) %>%
  units::drop_units()

# assert that a tract must be at least 1/10 mile from a park
distances <- pmax(distances, 0.1)
```

```{r write}
write_rds(distances, "data/distances.rds")
write_rds(W, "data/W.rds")
```

## Logsums
Here's the function we use to calculate the logsums
```{r}
#' Calculate destination choice logsums from a distance matrix
#'
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients.
#'   If we submit two variables the tweets are ignored.
#'
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a
#'
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-0.4,0.2,0.001)){

  # A is n x p
  a <- betas[1] * d

  # B is p x 1
  b <- betas[2] * sizes

  if(!is.null(tweets))
    b <- b + betas[3] * tweets

  # calculate observed utility by adding the weighted park-level attributes
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(a, 2, b, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))

}
```