### 2019-07-15 02_parks_data.Rmd
Revise open spaces builder
Calculate trails, playgrounds, courts, etc.

---
title: "Parks Data"
author: "Greg Macfarlane"
date: "6/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(leaflet)
library(tidyverse)
library(RSocrata)
library(leaflet)
library(sf)
```

In this document we get the parks data from the New York City open data
repository and do some cleaning. First, we get the parks dataset. The original
data dictionary is available at 
[on Github](https://github.com/CityOfNewYork/nyc-planimetrics/blob/master/Capture_Rules.md#open-space)
```{r parks}
parks_url <- "https://data.cityofnewyork.us/api/geospatial/y6ja-fw4f?method=export&format=GeoJSON"
parks <- st_read(parks_url, stringsAsFactors = FALSE) %>%
  mutate(shape_area = as.numeric(shape_area))
```

The New York City open data portal also contains a list of designated trails 
within parks.

```{r trails}
trails <- read_csv("https://data.cityofnewyork.us/api/views/vjbm-hsyr/rows.csv?accessType=DOWNLOAD")
trails2 <- jsonlite::fromJSON("https://www.nycgovparks.org/bigapps/DPR_Hiking_001.json")
parks_with_trails <- c(trails$ParkID, trails2$Prop_ID)
```

## Cleaning
The map of Flushing Meadows in Queens below shows some of the difficulties of this dataset:
  - First, the park is constructed of multiple outlines that limits the size of 
  the total park. This makes Flushing Meadows appear smaller than Central Park, 
  which is a single large park. 
  - Second, the dataset includes tennis parks and baseball fields as separate 
  facilities. We want to include these as attributes of their containing park
  rather than independent facilities.
  - Third, parking lots for stadia (in this case Citi Field) are included as 
  open park space. We should remove these as they are categorically unlike other
  green spaces.
  - Fourth, important green spaces like cemeteries are in a different dataset.

```{r fm1}
leaflet(st_buffer(parks %>% filter(grepl("Flushing Meadows", park_name)), dist = 0.000001)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(label = ~as.character(park_name), stroke = 0.5)
```

First, we manually remove the park facilities corresponding to stadia and their
associated parking lots, and then tag facilities that are not explicitly
park boundaries.

```{r notparks}
notparks <- c(9498000833, 14498000872, 14498000794, # citi field, Yankee stadium, garage
              9498000631)  # USTA Billie Jean Tennis Center

parks <- parks %>% 
  filter(!(source_id %in% notparks)) %>%
  mutate(
    courts = ifelse(grepl("Courts", landuse), TRUE, FALSE),
    playgrounds = ifelse(grepl("Playground", landuse), TRUE, FALSE),
    trails = ifelse(parknum %in% parks_with_trails, TRUE, FALSE)
  ) 
```

With those areas removed, we build a 30-foot buffer around each park and 
dissolve interior boundaries based on the park number. Tennis courts, etc. become
boolean variables indicating whether the park has the facilities or not. This
will make parks that are effectively the same facility a single polygon feature.
We also re-calculate the area of the parks in acres and remove parks smaller
than half an acre.

```{r dissolve}
dissolved <- parks %>%
  filter(!is.na(parknum)) %>%
  group_by(parknum) %>%
  filter(shape_area > 0.5 * 43560 | courts) %>%
  summarise(
    area = sum(shape_area), 
    courts = any(courts), 
    playgrounds = any(playgrounds),
    trails = any(trails),
    park_name = park_name[1]
  ) %>%
  mutate(sqft = st_area(.), acres = units::set_units(sqft, acres)) %>%
  st_transform(3628)
```

```{r fm3}
leaflet(st_transform(dissolved, 4326) ) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5,)
```

### Cemeteries
Some important green spaces that can be used for physical activity but are not
parks include cemeteries. These are stored in a separate dataset called Open
Space (Other) with the feature code `2500`.

```{r cemeteries}
cemeteries_url <- "https://data.cityofnewyork.us/api/geospatial/pckb-8r2z?method=export&format=GeoJSON"
cemeteries <- st_read(cemeteries_url) %>%
  filter(feature_co == 2500)
```

```{r d_cemeteries}
dissolved_cemeteries <- cemeteries %>%
  st_transform(3628) %>%
  st_buffer(5) %>%
  mutate(sqft = st_area(.), acres = units::set_units(sqft, acres))
```

```{r cemeteries_map}
leaflet(st_transform(dissolved_cemeteries, 4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5)
```

Now we can join the cemeteries data to the parks data and write out the results
```{r write}
rbind(
  dissolved %>% 
    transmute(id = parknum, name = park_name, acres, courts, playgrounds, trails, 
              type = "park"),
  dissolved_cemeteries %>% 
    transmute(id = source_id, name, acres, courts = FALSE, playgrounds = FALSE, 
              trails = FALSE, type = "cemetery")
  ) %>%
  filter(as.numeric(acres) > 0.5) %>%
  st_simplify(preserveTopology = TRUE) %>%
  write_rds("data/open_spaces.rds") %>%
  geojsonio::geojson_write(file = "data/open_spaces.geojson")
```