### 2019-07-15 03_logsums.Rmd
Revise logsum function
It's now more simple with a basic impedance and size term.

---
title: "Logsums"
author: "Greg Macfarlane"
date: "6/7/2019"
output: html_document
---

```{r setup}
library(tidyverse)
library(sf)
library(spdep)
library(leaflet)
```

In this document we calculate the distance between each tract in New York City 
and each park to help calculate a logsum-based measure of accessibility. We 
cleaned and gathered the data in the previous two pages.

```{r load}
open_spaces <- read_rds("data/open_spaces.rds")
tracts <- read_rds("data/tracts.rds")
```

We are going to do a point-to-polygon distance calculation. The geometric centroid
of the tracts are not satisfying, so we will use the population-weighted centroid
instead.

```{r pop_weighted_centroids}
pop_weighted_centroids <- tracts %>% tbl_df() %>%
  dplyr::select(geoid) %>%
  left_join(
    read_csv("https://www2.census.gov/geo/docs/reference/cenpop2010/tract/CenPop2010_Mean_TR36.txt") %>%
      transmute(geoid = str_c(STATEFP, COUNTYFP, TRACTCE), LATITUDE, LONGITUDE)
  ) %>%
  # convert to sf and project into the feet-based projection used in the
  # parks dataset.
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(3628)
```

The population-weighted centroid also serves as the orientation for the spatial
weights matrix.

```{r W}
# determine which tracts are neighbors of each other. Need to be less than
# 1.8 miles away from the centroids
tracts.dnn <- dnearneigh(as(pop_weighted_centroids, "Spatial"), 0, 1.8 * 5280)
# calculate the inverse distance to use as a weight on the neighbor connection
dists <- nbdists(tracts.dnn, as(pop_weighted_centroids, "Spatial"))
dists.inv <- lapply(dists, function(x) 1 / x)
# get the spatial weights matrix and its trace to use in any impact analyses
W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv,
              zero.policy = TRUE, style = "W")
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
print(W, zero.policy = TRUE)
```

## Distances
We want to make sure that the distance between a park and a tract centroid is
based on the edge of the polygon instead of the centroid thereof, so that spaces
at the corners of Central park, for instance, don't appear as longer than they should be.
It seems that things are working appropriately.

```{r test_measurement}
central_park <- open_spaces %>% filter(name == "Central Park")
pop_weighted_centroids$central_park <- log(as.numeric(st_distance(pop_weighted_centroids, central_park)))

dmeasure <- colorNumeric(palette = "Blues", domain = pop_weighted_centroids$central_park)
leaflet(pop_weighted_centroids %>% st_transform(4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addCircleMarkers(color = ~dmeasure(central_park))
```

Now we can calculate everything.
```{r distances}
# distances from all tract centroids to all parks
distances <- st_distance(pop_weighted_centroids, open_spaces, byid = TRUE) %>%
  units::set_units(miles) %>%
  units::drop_units()

# assert that a tract must be at least 1/10 mile from a park
distances <- pmax(distances, 0.1)
```

```{r write}
write_rds(distances, "data/distances.rds")
write_rds(W, "data/W.rds")
```

## Logsums

Here's the function we use to calculate the logsums. 
```{r}
#' Calculate destination choice logsums from a distance matrix
#'
#' @param impedance An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param size_term A p-length vector of the park size terms
#'
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a
#'
calculate_park_logsums <- function(impedance, size_term){

  # calculate observed utility by adding the weighted park-level attributes
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(impedance, 2, size_term, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))
}
```

```{r gravity}
logsum_vars <- tracts %>%
  transmute(
    geoid, 
    # Log-sum based utilities, without and with Twitter data
    access_ls = calculate_park_logsums(
      log(distances), open_spaces$acres, betas = c(-0.5, 0.05)),
    # Normalize the logsum onto the Z-scale
    access_ls_n = (access_ls - mean(access_ls) ) / sd(access_ls)
  ) %>%
  st_set_geometry(NULL)
```

```{r binary}
tracts_access <- tracts %>%
  left_join(logsum_vars, by = "geoid") %>%
  mutate(access_025 = ifelse(geoid %in% inside_025$geoid, TRUE, FALSE))

tracts_sp <- as(tracts_access, "Spatial")
```

```{r}
# this interactive map is much easier to create and explore
pal <- colorQuantile("GnBu", tracts_access$access_ls_n, n = 5)
bin <- colorFactor("Dark2", tracts_access$access_025)
leaflet(tracts_access) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = "access", color = ~ pal(access_ls_n),
              label = ~as.character(round(access_ls_n,2))) %>%
  # addPolygons(group = "tweets", color = ~ pal(tweets_ls),
  #             label = ~as.character(round(tweets_ls,2))) %>%
  addPolygons(group = "1/4 mile", color = ~ bin(access_025)) %>%
  addLayersControl(baseGroups = c("access", "tweets", "1/4 mile"))
```