### 2019-10-04 manuscript/manuscript.Rmd
NEW DRAFT

---
title: >
  Modeling the impacts of park access on health outcomes: a utility-based 
  accessibility approach
author:
  - name: Gregory Macfarlane
    email: gregmacfarlane@byu.edu
    affiliation: Brigham Young University
    corresponding: gregmacfarlane@byu.edu
  - name: Nico Boyd
    email:
    affiliation: Fehr and Peers 
  - name: John E. Taylor
    email: jetaylor@gatech.edu
    affiliation: Georgia Institute of Technology
  - name: Kari Watkins
    email: kariwatkins@gatech.edu
    affiliation: Georgia Institute of Technology
address:
  - code: Brigham Young University
    address: Civil and Environmental Engineering Department, Brigham Young University 430 Engineering Building, Provo, UT, 84602
  - code: Fehr and Peers
    address: Los Angeles, California
  - code: Georgia Institute of Technology
    address: School of Civil and Environmental Engineering, Georgia Institute of Technology, 755 Atlantic Drive, Atlanta, GA, 30332
abstract: >
  Recent research has underscored the potential for public green spaces to
  influence individual and societal health outcomes, but empirical measurements
  of such influences have yielded mixed results to date, with particular
  disagreement surrounding how access to parks ought to be defined while
  controlling for alternate explanations. In this paper we apply a comprehensive
  measure of park accessibility drawn from random utility choice theory, which
  avoids arbitrary assertions of proximity while incorporating potentially
  numerous amenities and attributes of both the parks and the population. We
  apply this utility-based accessibility measure to correlate Census tract-level
  obesity and physical activity rate estimates from the Centers for Disease
  Control and Prevention's 500 Cities project with tract-level American
  Community Survey socioeconomic data in New York City, paired with geographic
  open space data from New York City. Controlling for the
  socioeconomic variables and spatially correlated error terms, we show a
  positive and significant relationship between park access and physical
  activity rates. The data also suggest a negative relationship between park
  access and obesity rates, beyond what is expected through physical activity
  and socioeconomics.
  In doing so, this research contributes a more
  comprehensive modeling approach for measuring the impact of park access on
  health, and may improve our understanding of the role  parks and access to
  them can serve in furthering public health objectives.
author_summary: >
  Greg Macfarlane is an assistant professor of Civil Engineering at Brigham
  Young University in Provo, Utah.

output: 
  rticles::elsevier_article:
    includes: 
      in_header: preamble.tex
csl: elsevier-harvard.csl
bibliography: library.bib
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
library(texreg)
library(here)

options(tigris_use_cache = TRUE)
opts_chunk$set(cache=FALSE, message = FALSE, echo = FALSE, warning = FALSE)
library(leaflet)

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",", digits = 4, drop0trailing = TRUE)
})

```

# Introduction

The United States and other developed nations face an epidemic of obesity and
chronic diseases, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer. These diseseases
can severely limit the lifespan of affected individuals [@WHO2014], with 
substantial financial costs for treatment borne both privately and socially
[@Finkelstein2009]. Though a moderate amount of regular physical activity
has been established as an effective strategy for reducing and managing obesity
and many of the aforementioned chronic diseases [@CDC2009; @Durstine2013],
a large portion of U.S. adults do not participate in sufficient physical
activity [@Wolf2008].

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise.
Over the last several years a new *social* model of health has evolved,
described by @Duhl1999 
``as an outcome of the effects of socioeconomic status, culture,
environmental conditions, housing, employment and community influences''[-@Duhl1999, p.7].
In this new paradigm, resources provided via civil infrastructure --- in
particular parks and other public green spaces --- play a crucial role in
promoting and sustaining public health [@Bedimo-Rung2005; @Wells2007; @Coutts2008].
After all, it does little to lecture individuals on the importance of exercise
if their community is built in such a way that such exercise is expensive,
unenjoyable, or unsafe.

In spite of the potential for green spaces to influence public health outcomes 
and numerous studies attempting to empirically estimate the strength of these
outcomes, evidence to date has been mixed. A major challenge researchers have
faced is a proliferation of techniques for measuring access to green spaces, of
which many are plausible but lack robust theoretical basis. In this
paper, we apply a holistic and flexible measurement for park accessibility
based in activity location choice theory. This utility-based accessibility measure
compares the continuous distance to all parks in the region, weighted against
the sizes of each park and its other assorted amenities.

We apply this measurement to study the link between park accessibility and
attractiveness and Census tract-level aggregate physical activity participation
and obesity rates in New York City, controlling for spatially correlated
unobserved effects, spatial spillovers of covariates, and tract-level
socioeconomic characteristics. We also demonstrate how the utility-based
accessibility metric can potentially be expanded to account for other measures
of a park's attractiveness including its observed social media activity and the
presence of various park amenities. The paper concludes with a discussion of
opportunities for future research.

# Literature Review 
\label{sec:litreview}

The large majority of extant research on the topic of park access and public health
outcomes has been involved comparing metropolitan regions against each other. For example,
@West2012 used park data from
the Trust for Public Land’s 2010 "City Park Facts" and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine the relationship
between the density of park land and physical activity and obesity rates for 67
metropolitan statistical areas in the U.S. The findings in this case conformed to 
expectation; that is, the study found a significant positive association
between park density and physical activity and a significant negative
association between park density and obesity.
@Larson2016 similarly used self-reported scores on the Gallup-Healthways
Wellbeing Index to evaluate the relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities. While the authors found
positive associations between wellbeing and park quality and accessibility,
these relationships were not statistically significant.
Conversely, @Richardson2012 examined the relationship between total urban
greenspace and mortality rates for selected maladies; though the authors did not
find a statistically significant relationship between the quantity of urban
greenspace and mortality caused by lung cancer, diabetes, heart disease or car
accidents, they did find that all-cause mortality was --- oddly --- *higher* in greener
cities. Finally, in a meta-analysis of 20 peer reviewed journal articles exploring the
relationship between parks and objectively measured physical activity,
@Bancroft2015 found that five studies reported a significant positive
association between the two, six studies produced mixed results, and nine
studies found no association at all.

Metropolitan-level analyses such as those described above do not capture the
within-city variation in accessibility to parks that exist in many cities. Some
cities with large amounts of total greenspace may nevertheless unequally
distribute this space throughout the city, leading to areas with poor access.
Conversely, a city with a smaller overall proportion of greenspace may give all
of its citizens better access to sufficient greenspace to meet their needs or
wants for physical activity. Considering access at the neighborhood level within
a single city may also eliminate some regional or cultural fixed effects
affecting metropolitan-level analyses.

## Park Accessibility Measures

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there. Accessibility is an abstract
concept with tempting quantifiability [@Handy1997], perhaps explaining the proliferation of
calculation mechanisms. @Dong2006 present a helpful mathematical heirarchy of
common mechanisms, which we follow here. Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_{ij}$ is 
within an "isochrone," or a geometric buffer defined by some threshold $D$,

\begin{equation}\label{eq:isochrone}
A_i = \begin{cases} 
      1 & d_{ij}\leq D\\
      0 & d_{ij} > D
   \end{cases}
\end{equation}

Variations of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and without this threshold. In other applications $D$ might be more
variably defined, such as the presence or percent of green space within a
political or statistical boundary [@Mitchell2008; @Stark2014] or within a specific distance
threshold [@Kaczynski2014]. It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers in its application can inhibit
holistic analysis [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of point $i$ is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_{ij})$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function and
$S_j = 1$.
@Giles-Corti2005 compare various gravity-based accessibility scores with an
isochrone specification and show that the former are more predictive of walking
behavior; that is, the attributes of a park matter influence walking more than
merely the distance to it. @Zhang2011 develop a population-weighted,
gravity-based accessibility to parks metric with a national scope, though they
do not examine its correlation with health outcomes.

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply
the multinomial logit model [@McFadden1974], the expected consumer surplus
enjoyed by this individual can be shown to be
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities*, $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$ [@Bruce1977]. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utilityV}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation}
The coefficients $\vect{\beta}$ are frequently estimated from household
surveys, though in the absence of a survey we may assert reasonable values.

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\beta + \beta_d d_{ij}) = X_j f(d_{ij})$
[@Daly1982]. The distinction between gravity and utility-based specification is
meaningful, however. Primarily, it becomes possible to construct accessibility
statistics based on revealed behavioral preferences rather than calibrated or
asserted values [@Handy1997]. @Kaczynski2016 present what is effectively a 
utility-based accessibility score they call "ParkIndex", with the explicit
motivation of developing a uniformly applicable park accessibility statistic,
though they limit this statistic as a way to add heterogeneity within an
isochrone analysis.

Both gravity and utility-based specifications hold several advantages relative
to isochrone-based accessibility metrics more commonly found in the literature.
First, all individuals are defined as having some access to all parks, rather
than an arbitrary cutoff asserted by the researcher. This allows for the fact
that some people are more or less sensitive to distances, and that distance is a
continuous, non-binary phenomenon. It defies reason to assume a person living
11 minutes via walking from a park always has meaningfully different
accessibility than someone living 9 minutes away. Second, the random utility
formulation allows the
researcher to include -- in principle -- any attribute of the park as part of
its utility specification. This suggests that not all parks are equal, and that
a large park with many amenities such as Central Park in Manhattan may provide
health and activity benefits over a much larger area than a smaller community
square.

In spite of its flexibility and basis in choice theory, utility-based
accessibility measures have not received much application in the accessibility
literature compared with distance-based or even gravity-based measures.
@Vale2016 present a descriptive classification of active accessibility techniques, and
explicitly dismiss utility-based metrics for incorporating randomness. The
accessibility formula presented in \ref{eq:utility} is the expectation of a
random utility process, and is not in and of itself random any more than the
gravity model is random.
Utility-based accessibility metrics are commonly used, however, in alternatives
analyses of transit infrastructure improvements [@DeJong2007]. A reason for this
is likely that a regional travel demand model is available to the analysts, thus
making calibrated and multimodal logsums readily available [@Geurs2010].

# Empirical Application
In this section, we develop a model with data for New York City, where
we compute a set of utility-based accessibility to parks scores for each tract
and model the relationship between this measure and physical activity rates,
controlling for spatial effects and socioeconomic factors. We subsequently model
the effect of the accessibility metric on obesity rates, accounting for physical
activity and the other controls.

## Data

```{r tracts}
# built in 01_health_data.Rmd and 03_logsums.Rmd
tracts <- read_rds(here("data/tracts_access.rds"))
```

This study uses data available to the public from a variety of federal and state
data agencies^[The datasets, as well as the analysis code, are available on GitHub
at \url{https://github.com/gregmacfarlane/parks_access}.].
The Centers for Disease Control and Prevention makes small-area estimates on
key health indicators available through its 500 Cities data program [@CDC5002016].
The indicators are multilevel aggregations and imputations of BRFSS responses
[@Wang2018; @Wang2017], and have been recently used to study the tract-level
link between gentrification and urban health [@Gibbons2018]. We use two
indicators as our dependent variables in this study: the share of adults
in a Census tract who are obese, and the share of adults who participate in no
leisure-time physical activity. To improve clarity in our interpretation, we use
the complement of the second variable --- the share of tract adults who
participate in *some* physical activity --- even if the amount may not be
sufficient to affect overall health. Both indicators are obtained for the year
2016.

To the health data, we join socioeconomic data collected through the Census
Bureau API via the \texttt{tidycensus} package for R [@Walker2019]. The
primary dataset is a geographic polygons layer of Census tracts in the five
boroughs of New York City. We append to each Census tract relevant
socioeconomic data for each tract from the American Community Survey
2013-2017 5-year estimates. For a small handful of tracts in our sample, Census
supressed the median income estimate; these appear to be primarily wealthy
tracts and in almost all cases the CDC estimates of obesity and physical
activity are missing as well. After removing these tracts from the estimation
dataset, we have `r nrow(tracts)` complete cases. Table \ref{tab:tractsdata}
presents key descriptive statistics for these data.

```{r parks}
# This data is created in 02_parks_data.Rmd
parks <- read_rds(here("data/open_spaces.rds"))
```

In a destination choice framework, the tracts represent the "origins" and the
"destinations" are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these distinct sub-facilities, we created 
variables for each containing park indicating the presence of sports courts,
playgrounds, and trails.

To this dataset of parks and regularized green spaces, we add a polygons layer of
cemeteries in New York that are open to the public. Cemeteries are important
green spaces that can be used for many types of physical activity. These
operations leave us with `r nrow(parks)` distinct green spaces; for simplicity 
we will refer to both parks and cemeteries as "parks" going forward. 
For each park and 
we calculate the size of the park in acres. Using a Twitter application
programming interface (API) implemented with the Python package Tweepy,  we also
collected and stored tweets containing geotags of precise geographic coordinates
located within the boundaries of the parks; from this information,
we were able to segregate tweets that were generated within a park in September
2014. This geolocated twitter activity could provide an additional point of information
on the degree to which individual parks are actually used [@Wang2016].

```{r tractsdata, results="asis", message=FALSE, warning=FALSE, cache=FALSE}
# This is an index pairing the variable names with the order we want them
# to be shown in in the table.
tracts_data_index <- tibble(
  variable =  c(
    "obesity", "physact", "income",
    "density", "fulltime", "college", "single",
    "youth", "young_adults",  "seniors", "black", "asian", "hispanic",
    "other"),
  source = c("CDC 500 Cities", "CDC 500 Cities",
             rep("ACS", 12))
) %>%
  mutate(order = row_number())

# Get the tract SE / health data, calculate group statistics, and get it
# in the right order
tract_summary <- tracts %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  dplyr::select(tracts_data_index$variable) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    min = min(value),
    `msd` = paste(median_iqr(value)),
    max = max(value)
  ) %>%
  inner_join(tracts_data_index) %>%
  arrange(order) %>%
  mutate(
    variable = c(
      "Obesity: Share of adults over 18 who are obese",
      "Physical Activity: Share of adults over 18 who engage in some leisure-time physical activity",
      "Income: Median tract income",
      "Density: Households per square kilometer",
      "Fulltime: Share of adults over 18 with full-time work",
      "College: Share of adults over 24 with a college degree",
      "Single: Share of adults over 18 living alone or in a non-partnership household",
      "Youth: Share of population under 18",
      "Young adults: Share of population between 18 and 34",
      "Seniors: Share of population who are 65 and over",
      "Black: Share of population who is black",
      "Asian: Share of population who is Asian",
      "Hispanic: Share of population who is Hispanic",
      "Other: Share of population who belong to other minority groups")
  ) %>%
  dplyr::select(-order)

tract_rows <- nrow(tract_summary)

park_summary <- parks %>%
  st_set_geometry(NULL) %>%
  # TODO: This should be tweets
  dplyr::select(acres, courts, playgrounds, trails, tweet_count) %>% 
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    min = min(value),
    `msd` = paste(median_iqr(value)),
    max = max(value)
  ) %>%
  mutate(
    variable = c("Size: Park size in acres",
                 "Courts: Presence of sport courts / ball fields",
                 "Playgrounds: Presence of playgrounds",
                 "Trails: Presence of trails",
                 "Tweets: Tweets emanating from park in September 2014"),
    source = c(rep("NYC", 4), "Twitter API")
    )

bind_rows(tract_summary, park_summary) %>%
  separate(variable, c("var", "desc"), sep = ": ") %>%
  mutate(
    msd = str_replace(msd, " \\(", "\\\\\\\\ \\("),
    msd = str_c("\\makecell{", msd, "}")
  ) %>%
  kable(escape = FALSE, digits = 2, format = "latex",
        table.envir = "table*", booktabs = TRUE,
        caption = "Descriptive Statistics of Tract and Park Variables",
        col.names = c(" ", "Description", "Minimum", "Median (IQR)", "Maximum", "Source")) %>%
  kable_styling(latex_options = "scale_down") %>%
  column_spec(2, width = "2in") %>%
  group_rows(paste("Tract Variables, N =", nrow(tracts)), 1, tract_rows) %>%
  group_rows(paste("Park Variables, N =", nrow(parks)), tract_rows + 1, tract_rows + nrow(park_summary))
```

## Accessibility 
\label{subsec:accessibility}

We calculate a set of utility-based accessibility statistics for each tract
in New York City. The most basic utility specification includes only the
the park size in acres and the distance from the population-weighted tract centroid
to the boundary of the park in miles,
\begin{equation}\label{eq:u_access}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij})
\end{equation}
The logarithmic transform allows for diminishing marginal utility of distance
and park size: A 1-mile increase to a trip matters more for a 1-mile trip than a
10-mile trip. @Macfarlane2019 estimated destination choice 
parameters for park trips in the Alameda County (Oakland), California using
this utility specification, obtaining values of $\lambda_s = 0.373$ and 
$\lambda_d = -1.98$. The ratio of these estimates implies people are willing to
travel roughly six times further to reach a park twice as large. In the absence 
of park trip destination choice coefficients in New York City, we adopt these
previous estimates.

@Kinnell2006 surveyed park users in New Jersey and applied a multinomial logit
model to determine which factors influenced its perceived utility as a park
trip destination. By transferring estimates of common variables between our
data and this survey-based model, we can create a second utility specification
as
\begin{equation}\label{eq:u_multi}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_t * trails + \\
  \lambda_{c} * courts + \lambda_p * playgrounds
\end{equation}
with $\lambda_t = 0.99$, $\lambda_c = 0.43$, and $\lambda_p = 0.26$. Kinnell et al.\
did not transform their size and distance estimates, so we retain the previously 
estimated and applied values. We assume that the other covariates in the 
model which we do not have available to us (e.g., boat launches, picnic areas) are
orthogonal to the included parameters and leaving them out will not affect the
values of the included coefficients.

The third utility specification includes the number of geolocated tweets
emanating from within the park in September of 2014 in addition to the
basic specification in Equation \ref{eq:u_access},
\begin{equation}\label{eq:u_tweets}
V_{ij} = \lambda_s * \log(size_j) + \lambda_d * \log(distance_{ij}) + \lambda_{tw} * YJ(tweets)
\end{equation}
with $YJ(x)$ being a Yeo-Johnson transformation that implements a diminishing
marginal return similar to $log(x)$, but where $YJ(0) = 0$ [@Yeo2000]. In this
case we have no external information describing how destination choice to
parks may be affected by the twitter activity in the park, so we assert a 
value of $\lambda_{tw} = 0.1$

We then calculate the utility-based accessibility of each tract $A_i$ as defined
in Equation \ref{eq:utility} under each utility specification. Recall that the 
total value of the accessibility is relative to an unknown scalar $C$; for this 
reason we standardize the accessibility values for all tracts within each
utility specification, 
\begin{equation}\label{eq:v_logged}
A_i' = \frac{A_i - \bar{A}}{sd(A)}
\end{equation}

Figure \ref{fig:map} shows a map of New York City with each tract highlighted
based on its relative basic utility-based accessibility score (Equation
\ref{eq:u_access}). The most continuous region of good park access is in upper
Manhattan and the Bronx, bracketed by Central Park and the Bronx River.
Conversely, some of the poorest-accessibility areas are in Brooklyn tracts not
immediately adjacent to Prospect Park. Because the accessibility statistic is
normalized, the worst values are slightly below $-3$, the best somewhere above
$3$.

For comparative purposes, we also employ an isochrone analysis using the
10-minute walk threshold calculated for ParkScore [@parkscore2019]. If the
population-weighted centroid of a tract is within a 10-minute walk of a green space
as defined by ParkScore, the tract is considered as having "access" to a park.
The centroid is necessary as all tracts in New York City have at least *some* 
intersection with the 10-minute walk buffer; as it is, only 
`r sum(!tracts$walk_10)` of the `r nrow(tracts)` tract centroids are not located 
within this buffer.

```{r map, fig.env="figure*", echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.cap="Normalized utility-based park accessibility values in New York City."}
library(ggmap)
map <- ggmap(
  get_stamenmap(unname(st_bbox(tracts)), 
                maptype = "toner-lite"), extent = "device")
map +
  geom_sf(data = tracts, inherit.aes=FALSE,
          aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
          alpha = 0.8, color = NA) +
  scale_fill_brewer("Accessibility", palette = "GnBu")
```

## Model 
\label{subsec:model}

We predict either the physical activity or the obesity rate ($y$) in a census
tract as a linear function of the tract's sociodemographic characteristics $X$
and accessibility to parks $A$,
\begin{equation}\label{eq:themodel}
 y = X\beta + \beta_{a}A + \varepsilon
\end{equation}

As the observations are related to each other spatially, it is likely that this
process involves spatial spillovers of one kind or another. A complete treatment of
spatial econometrics is not warranted here; the reader is directed to
@LeSage2009 as well as @LeSage2014.
Suffice it to say that the presence of spatial data generating processes can
negatively affect econometric interpretation in a variety of ways. Each of these
process relies on spatial autoregression, where elements of some random
variable $x$ are spatially dependent on other elements, 
\begin{equation}\label{eq:autoregression}
 x = \rho W x + \varepsilon
\end{equation}
with the strength of dependence an estimated parameter $\rho$, the structure of 
spatial dependence governed by the asserted matrix $W$ (details of $W$ follow below),
and some random uncorrelated residual $\varepsilon$. A collection of models is
available to represent spatial autodependence in the model residuals, the 
independent variables, and the dependent variable.

The specific spatial model applied is both an econometric choice
--- as improperly specified spatial models can lead to inconsistent estimates of
model parameters and standard errors --- as well as a philosophical choice about
the likely structure of spillovers in the problem at hand. In case of physical
activity and obesity rates, we believe that the outcomes are locally dependent:
That is, a particular individual decides whether or not to participate in
physical activity independently of whether his or her neighbors
participate in physical activity. This does not mean that we think it impossible that
the socioeconomic variables of neighboring tracts do not play a role. Mathematically, we assert that the autodependence
relationship on the dependent variable $y = \rho W y + \ldots$ has 
$\rho = 0$ in the current case.

With this relationship ruled out, it remains a possibility that the 
model residuals are spatially dependent,
\begin{equation}\label{eq:SEM}
 y = X\beta + \beta_a A + u; u = \lambda W u + \varepsilon
\end{equation}
that the outcome is partially dependent on the socioeconomic variables in *neighboring*
tracts,
\begin{equation}\label{eq:SLX}
 y = X\beta + \beta_a A + W X \gamma + \varepsilon
\end{equation}
or a linear combination of both.
\begin{equation}\label{eq:SDEM}
 y = X\beta + \beta_a A + W X \gamma + u; u = \lambda W u + \varepsilon 
\end{equation}
These models are referred as the spatial error model (SEM), the spatial lag of
$X$ model (SLX), and the spatial Durbin error model (SDEM). @Pace2008 suggest
that a Hausman-style test of the estimates of $\beta$ derived from the OLS 
(Equation \ref{eq:themodel}) and SEM (Equation \ref{eq:SEM}) can identify whether
the estimates are consistent. If the $\beta$ estimates are dissimilar, both 
specifications are inappropriate as they contain untreated omitted variables.
Similarly, if the SLX and SDEM estimates are dissimilar it is an indication of 
missing variables in that specification. On the other hand if the estimates of
both pairs of models are similar and the residual correlation parameter $\lambda$
is non-zero, the model accounting for spatially dependent residuals will produce 
proper estimates of the model standard errors.

Note that in the lag-$X$ specifications (SLX and SDEM) we exclude the accessiblity 
component $A$, that is we do *not* consider that the accessibility
in a neighboring zone will have an effect on physical activity or obesity rates.
In a practical sense, this is because the accessibility term $A$ is itself spatially
determined.

### Spatial Weights

The spatial weights matrix $W$ is constructed of individual elements where
$w_{ij} = 0$ if observations $i$ and $j$ are spatially independent of each other,
and $w_{ij} > 0$ if $i$ and $j$ are spatially related in some way. 
@Dubin1998 presents details on constructing $W$, but in this study we
assume that tracts sharing at least one common border point are neighbors of
each other^[We also considered a distance-discounted weights matrix, which
produced similar results. This matrix is selected for simplicity.]. The elements
of $W$ are then row-standardized so that each observation has an equal total
influence from all its neighbors.

```{r weights, warning = FALSE}
# This are developed in 03_logsums.Rmd
distances <- read_rds(here("data/distances.rds"))
W <- read_rds(here("data/W.rds"))
trMC <- trW(as(W, "CsparseMatrix"), type="MC")
```

## Results

```{r results}
slx_vs_sdem <- read_rds(here("data/slx_vs_sdem.rds"))
pa_models <- read_rds(here("data/pa_models.rds"))
obesity_models <- read_rds(here("data/obesity_models.rds"))
```

```{r termnames}
base_termnames <- c(
  "(Intercept)", "log(Density)", "log(Income)", "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)", "Seniors (65+)", "Black population share",
  "Asian population share", "Hispanic population share", "Other Minorities"
)

names_for_terms <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1])
)

lambda <- "$\\lambda$: spatial correlation"
access_terms <- c("Logsum", "Tweets",  "Attributes", "10-min walk")
```

```{r slx_hausman}
slx_hausman <- function(slx, sdem) {
  b1 <- slx$coefficients
  b2 <- sdem$coefficients
  db <- b1 - b2

  V1 <- vcov(slx)[names(b1), names(b1) ]
  V2 <- vcov(sdem)[names(b2), names(b2) ]
  dV <- V2 - V1

  tibble(
    Hstat = t(db) %*% MASS::ginv(dV) %*% db, 
    "p-value" = pchisq(Hstat, df = Matrix::rankMatrix(dV), lower.tail = FALSE)
  )
}

slx_sdem_htest <- slx_hausman(slx_vs_sdem[[1]], slx_vs_sdem[[2]])
```

To determine which spatial effects specification was appropriate, we estimated
SLX and SDEM models of physical activity participation rates as a function of
tract covariates and  accessibility. These two models are presented in Table 
\ref{tab:slx-sdem}. With only a few exceptions, the SLX and SDEM coefficient
estimates for corresponding covariates lie within the 95\% confidence intervals
of the other model. A Hausman-style test [@Pace2008] on the coefficients 
assumes a null hypothesis that the coefficients are equivalent; this test 
produces $p$-value of only $`r slx_sdem_htest[2]`$. Though this statistic is
technically beneath a common hypothesis test threshold of $\alpha = 0.05$, we
fail to reject the null hypothesis that the coefficients are equivalent, as 
the differences are at no point practically different from each other. The consequences
of rejecting the null hypothesis in this case would be to suggest a model with 
spatial dependence of the dependent variable; as stated above, we feel that such
a relationship is unreasonable. As a consequence of these analyses, we adopt the
SDEM specification moving forward as the spatial lag of the errors is 
significant and the overal log likelihood of the SDEM is higher than the SLX 
specification.

```{r slx-sdem, results = "asis"}
texreg(slx_vs_sdem, single.row = TRUE, ci.force = TRUE,  digits = 4,
       custom.coef.names = c(names_for_terms, "Accessibility", lambda), 
       # Goodness of fit (LM and SP are different)
       custom.gof.rows = list(
         "Log Likelihood" = lapply(slx_vs_sdem, logLik) %>% unlist(),
         "Num. Obs." = lapply(slx_vs_sdem, function(s) length(s$residuals))
       ),
       include.rsquared = F, include.adjrs = F, include.aic = F, include.loglik = F,
       include.parameters = F, include.nobs = F,
       use.packages = FALSE, booktabs = TRUE,
       caption = "Comparison of SLX and SDEM Coefficients",
       label = "tab:slx-sdem",
       custom.note = "%stars $y = $ tract-level physical activity rate.",
       caption.above = TRUE)
```

The coefficients in the SDEM model are of two general types: the direct effect
resulting from a tract's own attributes and the indirect effect ($\gamma$)
resulting from the spatially-weighted attributes of its neighbors. For example, a
percentage point increase in the share of young adults in a tract increases the
expected physical activity rate in a tract by 
`r slx_vs_sdem[[2]]$coefficients["young_adults"]` percentage points, and a
percentage point increase in the share of young adults
*in neighboring tracts* increases the expected physical activity rate by an 
*additional* `r slx_vs_sdem[[2]]$coefficients["lag.young_adults"]` percentage
points. This opportunity to examine spatial spillovers of attributes is a key 
benefit of using a spatial model.

For the most part the direct coefficients are highly significant and of the
expected sign. Tracts with higher shares of full-time workers, college-educated
adults, young adults, and high-income households all show a greater share of
individuals engaging in regular physical activity. Conversely, tracts with
greater population density and a greater share of single adults, children,
seniors, minorities, and low income households all have lower modeled expected
rates of physical activity. The indirect coefficients are less clearly significant,
with only density, median tract income, some age groups, and some minority groups
showing an indirect effect.

Table \ref{tab:pa-models} present the estimated spatial lag and accessibility
coefficients for three models using different specifications of the utility
equation, as well as the ten-minute walk buffer calculated by ParkScore. 
Note that the model called "Size and Distance" is precisely the model
labeled "SDEM" in Table \ref{tab:slx-sdem}. The controlling covariates are
suppressed for clarity but do not change meaningfully across the four specifications.
Complete model estimates are available from the authors. Each of the three
utility-based specifications is significant at the $\alpha = 0.05$ level or lower,
and is of the hypothesized directionality. That is, residents living in tracts 
with increased utility-based accessiblity to green spaces show a significantly
higher physical activity participation rate. Taking the rough mean of the three
estimated utility-based coefficient, moving from the least-accessible tract ($A_i = -3$) to the
most-accessible tract ($A_i = 3$) is expected to raise the physical activity rate
by `r 6 * .2` percentage points. With the understanding of the full 95\% confidence
interval, we could suggest that the true effect of excellent versus poor park access
is anywhere between `r 6 * .05` and `r 6 * .35` percentage points.

The 10-minute walk buffer is considerably less useful at isolating the effects
of park "access" on physical activity rates. As so few tracts are located
outside of the buffer, the standard error of the estimated effect is large and implies
an effect on activity rates anywhere from a loss of `r 6 * .37` percentage points
to a gain of `r 6 * 1.5`.

```{r pa-models, results = "asis"}
texreg(pa_models[c(2:5)], digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Twitter", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "lambda" = "$\\lambda$: spatial correlation", 
         "access_ls" = "Size and Distance", 
         "multi_ls" = "Amenities", 
         "tweets_ls" = "Twitter Activity", 
         "walk_10TRUE" = "10-minute walk"),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:pa-models",
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```

We now consider the impacts of a model where the dependent variable is
the *obesity* rate in a tract, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility metrics.
Table \ref{tab:obesity-models} presents estimates of the effect of all three
utility-based accessibility measures as well as the 10-minute walk buffer. 
As before, the estimates of the other variables are available from the authors.
The correlation between physical activity rates and obesity is clear; for every
percentage point increase in physical activity rate, the expected obesity rate
drops by $0.43$ percentage points in all four accessiblity specifications.
The indirect effect of physical activity rates on obesity in neighboring zones
is inconsequential, supporting our methodological assertion that public health
variables are not spatially dependent, though the factors leading to them may be.

As far as the accessibility values are concerned, none of the three
utility-based specifications nor the ParkScore 10-minute walk buffer are
correlated with obesity rates at the traditional 95\% confidence interval. 
It is important to consider the implications of the full range of the confidence
interval, however [@Amrhein2019]. The results suggest that after controlling
for tract socioeconomic variables, physical activity rates, and spatially correlated
error terms, obesity rates in tracts with excellent park accessibility are expected 
to be between `r 6 * 0.15` percentage points lower and `r 6 * .0053` percentage 
points higher than tracts with poor park accessibility. This range of values 
suggests that accessibility to parks may have obesity-reducing effects beyond
the effect of accessibility on physical activity rates worth further study.

```{r obesity-models, results="asis"}
texreg(obesity_models[c(2:5)], digits = 4, ci.force = TRUE, 
       booktabs = TRUE, use.packages = FALSE, 
       custom.model.names = c("Size and Distance", "Twitter", "Amenities", "10-Minute Walk"),
       custom.coef.map = list(
         "physact" = "Physical Activity", 
         "lag.physact" = "$\\gamma$: Physical Activity",
         "lambda" = "$\\lambda$: spatial correlation", 
         "access_ls" = "Size and Distance", 
         "multi_ls" = "Amenities", 
         "tweets_ls" = "Twitter Activity", 
         "walk_10TRUE" = "10-minute walk"),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:obesity-models",
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```

# Limitations and Future Research Direction

We readily acknowledge limitations in this study. As in any study conducted
with areal data, we are at risk of falling victim to the ecological inference
fallacy, where aggregate statistics mask or contradict disaggregated or
individual-level trends. A large-sample survey of individuals in New York City,
including measured physical activities and obesity would always be preferable
to the tract-level data used in this study. An ideal survey to address the
question would incorporate both sets of questions: physical activity and health
data on one hand and park use (including which parks were used and how
frequently) on the other. As no such dataset exists to our knowledge, this
tract-level aggregate analysis with asserted and transferred utility
coefficients is the possibility that remains.

This paper applies a previously-presented but little-used accessibility
statistic that could, in theory, accommodate many attributes of the destination
parks as well as the people who might use them. As an illustration: the
park-going population could be separated into at least four delineated clusters,
each preferring different amenities of a park:
\begin{itemize}
 \item{runners and cyclists: long, interesting trail systems}
 \item{sports players: soccer fields, basketball courts, or baseball diamonds}
 \item{families with small children: water features and playgrounds}
 \item{casual users: water features, gardens, performances, etc.}
\end{itemize}
An analyst could then compute the utility-based accessibility for each cluster
with different utility values for each park's amenities, and obtain a measure of
a neighborhood's accessibility to park features that its residents most care
about. In this paper, we proceed only incrementally towards this by adding
Twitter activity and some selected park amenities as an element of each park's 
attractiveness above and beyond its size. Exploring detailed definitions of
an increased number of amenities on one hand, and market segmentation strategies
on the other, could provide a more nuanced understanding of the relationship
between park accessibility and health outcomes.

As a travel impedance measure, we used the Euclidean distance between each tract's
population-weighted centroid and the nearest point on the edge of a park's border.
Euclidean distances have well-rehearsed limitations regarding their fidelity with
the underlying infrastructure network, etc. Network-based distances can also
suffer from challenges when applied to multimodal problems; these challenges are
exacerbated when the non-highway mode share is high, as is likely when considering
access to parks. A better metric of travel impedance may be a mode choice model
logsum, which weights all alternative travel alternatives against each other. 
We leave this as a recommendation for future research.

We compared our utility-based accessibility to a well-constructed isochrone
that has attracted attention in the literature. That said, virtually all tracts
in New York City are within this specific isochrone. It is possible that this 
isochrone may be more discriminating of park access in other cities, or that a
different definition would yield different results in New York City. This
observation however reinforces our suggestion that a measure explicitly based in
the utility of access is warranted.

Finally, this study is primarily focused on the hypothesis that accessibility to
parks encourages physical activity, which in turn reduces obesity. There are a
multitude of other hypotheses that might be proposed and tested with the basic
methodology we have employed here, or competing explanations for the outcomes we
have observed. It is distinctly possible, for instance, that individuals who wish
to exercise regularly in parks choose to live near them. Given that the CDC models
informing the small-area obesity and physical activity estimates presumably include variables
likely to influence such preferences (income, etc.), our investigation cannot
isolate the preferences from the effect. Controlling for such a
self-selection effect would be necessary to isolate the exogenous
impacts of park access on obesity or other health outcomes. And regarding these
other variables: this study did not consider potential relationships between park
access and hospitalization rates, life expectancy, respiratory disease,
mental health, or any number of potential beneficial outcomes hypothesized
or explored in the existing literature. Exploring these connections and their
underlying mechanisms should be a priority as city planners and urban architects
attempt to improve the quality of life of urban residents in the future.

# Conclusion

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
accessibility and these metrics, previous literature has produced mixed
findings, perhaps owing to the range of variables modeled and the coarse spatial
scale of the analyses. Using New York City as a case study, we presented a
holistic and flexible measurement for park accessibility that compares the
continuous distance to all parks in the region, weighted against the size of the
park and its other amenities, with weights determinable through revealed
preferences. In terms of physical activity, we found a positive relationship
where the least park-accessible tracts have an expected physical activity
participation rate between `r 6 * .05` and `r 6 * .35` percentage points lower than the most
accessible tracts. We also found a suggestive though not sigificant relationship
between park accessiblity and expected obesity rates *in addition to* the effect
of physical activity participation.

In both cases a common, isochrone-based analysis estimated a statistically weak
relationship with greater uncertainty. Isochrone analyses are relatively common in
the literature, perhaps because of the widespread availability of and training
in GIS software. In spite of their widespread use, isochrones are relatively
limited in terms of both their theoretical underpinnings and their flexibility
to accommodate attributes of parks beyond their proximity. And even proximity
may not be adequately handled, as the isochrone threshold may be arbitrarily
asserted by the researcher. The model we apply in this paper extends
a more comprehensive and flexible approach
for measuring the impact of park access on health outcomes. Adopting
utility-based accessibilities of the kind used in this study will allow
researchers to encompass the full range of park amenities in their accessibility
analyses, and to separate the definition of access from the study of its effects.
This will in turn enable planners to consider how multiple attributes
of a park --- from its location to its size to its amenities and beyond ---
benefit the health of the community the park serves.

## Acknowledgements
This project was funded by the Speedwell Foundation. We are grateful to Rachel
Samuels for helping to assemble the geolocated Twitter data.

# References {#references .unnumbered}