### 2019-11-06 manuscript/appendix.Rmd
Add appendix

---
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = F)
library(here)
library(tidyverse)
library(spdep)
library(knitr)
library(kableExtra)
library(texreg)
```

## Appendix

In this appendix we present the complete estimation results for the models 
relating different definitions of access to physical activity (in Table
\ref{tab:pa-fullmodels}) and to obesity (in Table \ref{tab:ob-fullmodels}). In
each case we also present a base model with no accessibility statistics for
comparison.

```{r}
pa_models <- read_rds(here("data/pa_models.rds"))
obesity_models <- read_rds(here("data/obesity_models.rds"))
```

```{r}
base_termnames <- c(
  "(Intercept)", "log(Density)", "log(Income)", "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)", "Seniors (65+)", "Black population share",
  "Asian population share", "Hispanic population share", "Other Minorities"
)

names_for_terms <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1])
)

lambda <- "$\\lambda$: spatial correlation"
access_terms <- c("Size and Distance", "Tweets",  "Attributes", "10-min walk")

ob_names <- c(base_termnames, "Physical Activity Rate")
ob_names <- c(
  ob_names, str_c("$\\gamma$: ", ob_names[-1])
)

```

```{r pa-fullmodels, results = "asis"}
texreg(pa_models, digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.coef.names = c(names_for_terms, lambda, access_terms),
       custom.model.names = c("No Access", access_terms),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:pa-fullmodels",
       caption.above = TRUE, include.aic = F, include.lr = F,
       scalebox = 0.7,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```

```{r ob-fullmodels, results="asis"}
texreg(obesity_models, digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.coef.names = c(ob_names, lambda, access_terms),
       custom.model.names = c("No Access", "Size and Distance", "Twitter", "Amenities", "10-Minute Walk"),
       caption = "Estimated Effect of Accessibility on Physical Activity Rates",
       label = "tab:ob-fullmodels",
       caption.above = TRUE, include.aic = F, include.lr = F, scalebox = 0.7,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```