### 2020-01-29 presentations/mentalhealth/05_mentalhealth_slides.Rmd

---
title: "Park Access and Mental Health"
author: "Greg Macfarlane | Civil and Environmental Engineering"
date: "12/12/2019"
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
library(tidyverse)
library(leaflet)
library(sf)
library(spatialreg)
library(spdep)
library(VGAM)
library(here)
```

## Accessibility{.pause}

 - How close is my destination?
 - How many destinations can I get to?
 - What is the quality of the destinations I can access?

## Utility-based Accessibility
Choice theory: the total value of a choice set is
$$A_i = \ln\left(\sum_{j\in J}\exp(U_{ij})\right), \quad U_{ij} = \beta_s size_j + \beta_d distance_{ij} + \ldots$$

Where:

  - $U_{ij}$ is the *utility* of person $i$ choosing option $j$
  - $\beta$ represents preferences for an attribute
  - Any attributes can be included

# Mental Health and Parks in New York City

## CDC 500 Cities

  - Derived from the Behavioral Risk Factor Surveillance System (BRFSS)
  - Synthetic data / modeled with margin of error
  - Includes tract-level estimates of obesity rates, physical activity rates, etc. including mental health

<blockquote>
Respondents aged ≥18 years who report 14 or more days during the past 30 days
during which their mental health was not good.
</blockquote>

## Mental Health in NYC
```{r loadtracts, echo = FALSE}
tracts <- read_rds(here("data/tracts_access.rds")) %>%
  st_transform(4326)
tracts$mhealth <- read_rds(here("data/tracts.rds")) %>% pull(mhealth)
```

```{r mhealth-map, echo = FALSE, fig.width=10}
pal <- colorBin("viridis", n = 10, domain = tracts$mhealth)
leaflet(tracts) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5, color = ~pal(mhealth), fillOpacity = 0.6) %>%
  addLegend(pal = pal, values = ~mhealth)
```

## Park Access Data

Estimated values of utility from cellular phone data and from pre-existing
surveys of park use

## Park Access in NYC

```{r access_map, echo = FALSE, fig.width=10}
# this interactive map is much easier to create and explore
pal <- colorBin("GnBu", c(tracts$access_ls, tracts$multi_ls), n = 5)
leaflet(tracts) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(group = "Kinner", color = ~ pal(multi_ls),
              label = ~as.character(round(multi_ls, 2))) %>%
  addLegend(pal = pal, values = ~multi_ls, title = "Access Score")
```

## Preliminary Econometric Model

```{r mhealth_models, echo = FALSE, results = "asis", warning = FALSE}
mhealth_models <- read_rds(here("presentations/mentalhealth/models.rds"))
texreg::htmlreg(mhealth_models, digits = 4, single.row = TRUE,
                custom.coef.map = list(
                  "yj(density)" = "log(Density)",
                  "yj(income)"  = "log(Income)",
                  "physact" = "Physical Activity Rate",
                  "multi_ls" = "Access to Parks"
                ), include.aic = FALSE, include.nobs = FALSE,
                doctype = FALSE, caption = "", star.symbol = "\\*",
                custom.note = "%stars Standard errors in parentheses.")
```

Model also includes educational attainment, age categories, ethnic group
categories, and spatial spillovers.

## Next Steps

Get more detailed cellular phone, transportation, and land use data

  - Parks including amenities
  - Churches
  - Recreation centers
  - Mental health providers

Better mental health / suicide risk data?