### 2020-03-16 05_mental_health.Rmd

---
title: "Mental Health Analysis"
author: "Greg Macfarlane"
date: "12/6/2019"
output: html_document
---
```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(leaflet)
library(sf)
library(spatialreg)
library(spdep)
library(VGAM)
```

In this document, I explore whether the park accessibility scores we have
estimated previously are also correlated with mental health in the way that they
suggest a correlation with obesity.

## Mental Health Data
The CDC's *500 Cities* project contains modeled health outcome, prevention, and 
risk factor data at the census tract level for the 500 largest cities in the
United States. The modeled health outcomes include arthritis, asthma, diabetes, 
heart disease, etc. including mental health. This outcome is defined as:

> Respondents aged ≥18 years who report 14 or more days during the past 30 days
during which their mental health was not good.

The data are derived from the Behavioral Risk Factor Surveillance System (BRFSS)
surveys, which are then modeled to project estimates at small-scales. 
We obtained mental health data for tracts in New York City.

```{r loadtracts}
tracts <- read_rds("data/tracts_access.rds") %>%
  st_transform(4326)
tracts$mhealth <- read_rds("data/tracts.rds") %>% pull(mhealth)
```

The median population-weighted tract seems to have about 12% of people reporting 
poor mental health.

```{r mhealth-hist}
ggplot(tracts, aes(x = mhealth, weight = population)) +
  geom_histogram() + theme_bw() + 
  xlab("% in tract reporting 14 of last 30 days with poor mental health") + 
  ylab("Population-weighted tracts")
```

The places with the worst mental health appear to be concentrated in the Bronx
and in Brooklyn, and the best in Manhattan's upper east side. At first glance,
This would suggest it is heavily correlated with income.

```{r mhealth-map}
pal <- colorBin("viridis", n = 10, domain = tracts$mhealth)
leaflet(tracts) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5, color = ~pal(mhealth), fillOpacity = 0.6) %>%
  addLegend(pal = pal, values = ~mhealth)
```

## Accessibility

We can also calculate a utility-based accessibility score for each tract to the
parks in the region. This score includes the number of parks, the size of them,
whether the parks have trails, playgrounds, and ball fields / courts, and the
euclidean distance to the park from the population-weighted tract centroid.

```{r access_map}
# this interactive map is much easier to create and explore
pal <- colorBin("viridis", c(tracts$access_ls, tracts$multi_ls), n = 5)
leaflet(tracts) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(group = "Kinner", color = ~ pal(multi_ls),
              label = ~as.character(round(multi_ls, 2)), 
              stroke = FALSE, fillOpacity = 0.7) %>%
  addLegend(position = "topleft", pal = pal, values = ~multi_ls, title = "Access Score")
```

## Spatial model
We now estimate a SDEM model where mental health is a response to tract income,
ethnic makeup, age, and other factors. The SDEM model considers spatial spillovers
in the error term and in the independent variables but not in the depdendent variable,

$$ y = X\beta + WX\gamma + u, u = \lambda W u + \varepsilon$$

```{r W}
W <- read_rds("data/W.rds")
models <- read_rds("data/obesity_models.rds")

contiglagged_vars <- tracts %>%
  st_set_geometry(NULL) %>%
  as_tibble() %>%
  transmute_at(
    vars(density, income, fulltime, college, single, youth, young_adults, seniors, black, asian, hispanic, other, physact),
    ~lag.listw(W, ., zero.policy = TRUE)
  ) %>%
  mutate(geoid = tracts$geoid) %>%
  dplyr::select(geoid, everything())

names(contiglagged_vars)[-1] <- str_c("lag.", names(contiglagged_vars)[-1])
lagged_tracts <- tracts %>% left_join(contiglagged_vars, by = "geoid")
```

```{r models}
yj <- function(x) VGAM::yeo.johnson(x, lambda = 0)
mhealth_base <- spatialreg::errorsarlm(
  mhealth ~ yj(density) + yj(income)  + fulltime + college + single + 
    youth + young_adults + seniors + black + asian + hispanic + 
    other +  yj(lag.density) + yj(lag.income) + lag.fulltime + 
    lag.college + lag.single + lag.youth + lag.young_adults + 
    lag.seniors + lag.black + lag.asian + lag.hispanic + lag.other + 
    physact,
  data = lagged_tracts, listw = W, zero.policy = TRUE)
mhealth_accs <- update(mhealth_base, formula = . ~ . + multi_ls)
mhealth_models <- list(base = mhealth_base, access = mhealth_accs)
```

```{r show_models}
texreg::htmlreg(mhealth_models, digits = 5)
```

```{r tidied}
tidy.sarlm <- function(x){
  s <- summary(x)
  df <- length(s$residuals) - length(s$coefficients) - 1
  tibble(
    term = gsub("[\\(\\)]", "\\.", rownames(s$Coef)),
    estimate = s$Coef[,1],
    std.error = s$Coef[,2],
    statistic = estimate / std.error,
    p.value = 2*(pt(abs(statistic), Inf, lower.tail = FALSE))
  )
}

tidied <- lapply(mhealth_models, broom::tidy) %>%
  bind_rows(.id = "model")

```

The model results are suggestive ($p$ value = 
`r filter(tidied, term == "multi_ls") %>% pull(p.value)`). This is the same
outcome we found for the obesity models. But it is worth noting that physical
activity rates are highly correlated with mental health, and access to parks
*is* a strong predictor of physical activity.

```{r dwplot}
weird <- scales::trans_new("signed_log",
                           transform=function(x) sign(x)*sqrt(abs(x)),
                           inverse=function(x) sign(x)*(abs(x)^2))
tidied %>%
  filter(term != ".Intercept.") %>%
  dotwhisker::dw_plot() + 
  theme_bw() + 
  scale_x_continuous(trans = weird)
```

### Utah Analysis

```{r}
tibble("Age" = c("18-25", "26-49", "50+"), "Serious MI" = c(7.5, 5.6, 2.7),
       "Any MI" = c(25.8, 22.2, 13.8)) %>%
  mutate(`Any MI` = `Any MI` - `Serious MI`) %>%
  gather("Type of Mental Illness", "Percent", -Age) %>%
  ggplot(aes(x = Age, y = Percent, fill = Malady)) +
  geom_bar(stat = "identity")
```

Get 500 cities mental health data for Utah County

```{r}
cdc_urls <- list(
  "mhealth" = "https://chronicdata.cdc.gov/resource/i2ek-k3pa.json?stateabbr=UT"
)

cdc500 <- lapply(cdc_urls, function(url){
  read.socrata(url, app_token = Sys.getenv("CDC_KEY"),
               stringsAsFactors = FALSE) %>%
    tbl_df() %>%
    filter(tractfips %in% acs$geoid) %>%
    transmute( geoid = tractfips, value = as.numeric(data_value ), year = year)
}) %>%
  bind_rows(.id = "measure") 
```

Get block groups
```{r}
bg <- tigris::block_groups("Utah", "Utah", class = "sf") %>%
  mutate(tract = substr(GEOID, 1, 11)) %>%
  left_join(cdc500, by = c("tract" = "geoid"))
```

map

```{r}
pal <- colorNumeric(palette = "magma", bg$value)
leaflet(bg) %>% addProviderTiles(providers$Esri.WorldGrayCanvas) %>% 
  addPolygons(color = ~pal(value))
```