### 2020-03-17 manuscript/appendix.Rmd
Remove Twitter analysis from manuscript
The models are still estimated and included in the website for future use and reference, but we can exclude them from the paper. Close #1.

---
output:
  pdf_document:
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = F)
library(here)
library(tidyverse)
library(spdep)
library(knitr)
library(kableExtra)
library(texreg)
```

## Appendix

In this appendix we present the complete estimation results for the models 
relating different definitions of access to physical activity (in Table
\ref{tab:pa-fullmodels}) and to obesity (in Table \ref{tab:ob-fullmodels}). In
each case we also present a base model with no accessibility statistics for
comparison.

```{r}
pa_models <- read_rds(here("data/pa_models.rds"))
pa_models$Tweets <- NULL
obesity_models <- read_rds(here("data/obesity_models.rds"))
obesity_models$Tweets <- NULL
```

```{r}
base_termnames <- c(
  "(Intercept)", "log(Density)", "log(Income)", "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)", "Seniors (65+)", "Black population share",
  "Asian population share", "Hispanic population share", "Other Minorities"
)

names_for_terms <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1])
)

lambda <- "$\\lambda$: spatial correlation"
access_terms <- c("Size and Distance", "Attributes", "10-min walk")

ob_names <- c(base_termnames, "Physical Activity Rate")
ob_names <- c(
  ob_names, str_c("$\\gamma$: ", ob_names[-1])
)

```

\begin{table*}
  \caption{\label{tab:pa-fullmodels} Estimated Effect of Accessibility on Physical Activity Rates}
  \begin{center}
  \scalebox{0.7}{
```{r pa-fullmodels, results = "asis"}
texreg(pa_models, digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.coef.names = c(names_for_terms, lambda, access_terms),
       custom.model.names = c("No Access", access_terms), table = F,
       caption.above = TRUE, include.aic = F, include.lr = F,
       custom.note = "%stars 95\\% confidence interval in brackets.")
```
}\end{center}
\end{table*}

\begin{table*}
  \caption{\label{tab:ob-fullmodels} Estimated Effect of Accessibility on Obesity Rates}
  \begin{center}
  \scalebox{0.7}{
```{r ob-fullmodels, results="asis"}
texreg(obesity_models, digits = 4, ci.force = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.coef.names = c(ob_names, lambda, access_terms),
       custom.model.names = c("No Access", access_terms), table = F,
       caption.above = TRUE, include.aic = F, include.lr = F, 
       custom.note = "%stars 95\\% confidence interval in brackets.")
```
}\end{center}
\end{table*}