### 2020-04-08 03_distances.Rmd
Find nearest node and shortest paths on network

---
title: "Distances"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
use_condaenv("ox", conda = "~/opt/anaconda3/condabin/conda")
```

We want to calculate network-based distances between blocks in New York City
and the edges of parks.  In this document, we do the following:

  - Obtain a highway network for New York City from OpenStreetMap
  - Determine coordinates for block group centroids and park edge nodes
  - Compute the shortest path between the centroid and the nearest edge node.

To do this, we are going to use the `osmnx` library for  python.

```{python setup}
import osmnx as ox
import networkx as nx

import numpy as np
import pandas as pd
import time
import multiprocessing as mp
```

## OpenStreetMap

We use the `osmnx` library to get a highway network graph for New York City from 
OpenStreetMap. We get the walk network because most people in New York City will
not be driving to get to their park.

```{python graph}
graph = ox.graph_from_place('New York, New York, USA', network_type='drive')
graph_proj = ox.project_graph(graph)
```

## Block Groups

We get a list of population-weighted block group centroids from Census.

```{python blockgroups}
# get the coordinates from census
url = "https://www2.census.gov/geo/docs/reference/cenpop2010/blkgrp/CenPop2010_Mean_BG36.txt"
blockgroups = pd.read_csv(url)

# pad the county and tract strings, and get the proper data types
blockgroups.STATEFP = blockgroups.STATEFP.apply(str)
blockgroups.BLKGRPCE = blockgroups.BLKGRPCE.apply(str)
blockgroups.COUNTYFP = blockgroups.COUNTYFP.apply('{0:0>3}'.format)
blockgroups.TRACTCE = blockgroups.TRACTCE.apply('{0:0>6}'.format)
blockgroups['GEOID'] = blockgroups[['STATEFP', 'COUNTYFP', 'TRACTCE', 'BLKGRPCE']].apply(lambda x: ''.join(x), axis = 1)

blockgroups['LATITUDE'] = blockgroups['LATITUDE'].astype(float)
blockgroups['LONGITUDE'] = blockgroups['LONGITUDE'].astype(float)

# filter to nyc
nyc_counties = ["081", "047", "061", "005", "085"]
blockgroups = blockgroups[blockgroups.COUNTYFP.isin(nyc_counties)]
blockgroups.set_index('GEOID', inplace=True)
```

## Open Spaces

We want to calculate distance from the block group centroid to the nearest
edge of each open space. The open spaces data created on the 
[previous page](02_parks_data.html) is a detailed simple features layer with a
polygon for each open space. The `osmnx` graph needs to have point coordinates 
to calculate network-based distances, however. To make this work, we

 1. simplify the polygon geometry to reduce the number of edge vertices. We do this
 with a tolerance of 100 feet.
 2. convert the polygon geometry into a boundary linestring
 3. sample points on the linestring with a density of 1 point every 500 feet. Parks
 with a permeter of less than 500 feet will have a single point.

```{r simplify_spaces}
# get open spaces file
open_spaces <- read_rds("data/open_spaces.rds") %>%
  select(id)  %>%
  # some polygons are way too detailed, so we want to simplify to 100 foot resolution
  st_simplify(dTolerance = 100, preserveTopology = TRUE)

# some parks have multiple polygons. We need these to be simpler individual features 
poly_from_multi <- open_spaces %>%
  filter(st_geometry_type(geometry) == "MULTIPOLYGON") %>%
  st_cast("POLYGON")

linestrings <- open_spaces %>%
  # join the original polygons back with the simplified multipolygons
  filter(st_geometry_type(geometry) != "MULTIPOLYGON") %>%
  rbind(poly_from_multi) %>%
  # cast the polygons to a linestring of the park perimeter
  st_cast("LINESTRING", group_or_split = TRUE)

point_samples <- linestrings %>%
  # sample points along the line, one point per 1000 feet.
  st_line_sample(density = 1/1000)

# append open space id
open_space_points <- tibble(id = linestrings$id, geometry = point_samples) %>%
  st_as_sf() %>%
  st_cast(to = "POINT")
```

The results of this data munging are given below. We have some issues where parks 
made of multiple polygons end up with internal points. Trying to get rid of these
by using complex hulls created other problems on linear parks. So we'll just go
with this.

```{r simplified-point-map}
leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  addCircleMarkers(data = open_space_points %>% st_transform(4326), radius = 0.001,
                   color = "red") %>%
  addPolygons(data = open_spaces %>% st_transform(4326))
```

Get the data into a form that we can pass it into python.

```{r park_points}
park_points <- open_space_points %>%
  st_transform(4326) %>%
  mutate(
    LONGITUDE = st_coordinates(.)[, 1],
    LATITUDE = st_coordinates(.)[,2]
  ) %>%
  st_set_geometry(NULL) %>%
  as_tibble()
```

## Distances

First, we bring the park points we sampled into a pandas data frame with
the park id as the index.
```{python park_points_py}
park_points = pd.DataFrame(r.park_points)
# TODO: run for all park points
park_ids = park_points.sample(5).id

park_points.set_index('id', inplace=True)
park_points = park_points.loc[park_ids]

```

We already created the block group centroids data frame earlier. Let's get the
nearest network nodes for each of the block groups and each park point.

```{python park_bg_nodes}
# TODO: run for all block groups
blockgroups = blockgroups.sample(5)

def findnode(df):
  df['node'] = df.apply(lambda row: 
    ox.get_nearest_node(graph, (row.LATITUDE, row.LONGITUDE)), axis=1)
  return df

def parallelize_dataframe(df, func, n_cores=4):
  df_split = np.array_split(df, n_cores)
  pool = Pool(n_cores)
  df = pd.concat(pool.map(findnode, df_split))
  pool.close()
  pool.join()
  return df

a = time.process_time()
findnode(blockgroups)
findnode(park_points)
b = time.process_time()
b - a
```

We want to create a data structure that we can quickly populate with the 
shortest path between a block group and the nearest park point. This is 
probably a two-level dictionary.

```{python blockgroup_dict}
# Create a dictionary for the block groups
blockgroup_dict = {}
for i in blockgroups.index.unique():
  blockgroup_dict[i] = {} # create a dictionary for the parks
  for j in park_points.index.unique():
    blockgroup_dict[i][j] = float("inf")

```

Now we want to loop through each block group and all of the park points to figure
out what the shortest path is.

```{python nearest}
# loop through block groups
a = time.process_time()
for bg in blockgroups.itertuples():

  # loop through parks
  for park in park_ids:
    these_points = park_points.loc[[park]]

    # loop through points
    for point in these_points.itertuples():
      # get length between park and this edge point
      length = nx.shortest_path_length(graph_proj, source = bg.node, target=point.node,
        weight = 'length')

      # check if this is the shortest one we've found; if so, update.
      if length < blockgroup_dict[bg.Index][park]:
        blockgroup_dict[bg.Index][park] = length

b = time.process_time()
b - a 
```