### 2020-04-20 04_accessibility.Rmd
Calculate logsum-based accessibilities

---
title: "Logsums"
author: "Greg Macfarlane"
date: "6/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(sf)
library(leaflet)
library(lubridate)
library(VGAM)
```

In this document we calculate the utility-based measure of accessibility, as well
as the 10-minute walk buffer.

## Distances
We cleaned and gathered the data parks data and data for tracts in the previous
two pages, as well as computed the network-based distance between each block
group and the nearest edge point of each park.

```{r load, message=FALSE}
open_spaces <- read_rds("data/open_spaces.rds")
blockgroups <- read_csv("data/blockgroups.csv")

# tigris not working, manual download
#bg_poly <- tigris::block_groups("New York", county =  c("081", "047", "061", "005", "085"))
bg_poly <- st_read("/Users/gregmacfarlane/Library/Caches/tigris/tl_2019_36_bg.shp") %>%
  filter(GEOID %in% blockgroups$GEOID) %>%
  transmute(geoid = GEOID) %>%
  st_transform(32618)
```

In calculating the distances, two things could happen. First, it is possible that
no path is found between the park and the block group centroid; this could be the
result of network miscodings or other errors in the network algorithm. We'll have
to take the results as they come, and we simply remove the park from that block
group's accessibility calculation. The other issue is when a block group centroid
and the park edge point match to the *same* node, resulting in a distance of zero.
Because we are going to log-transform the distance, we add ten meters to all paths.

```{r distance_adjust}
distance_files <- list.files("data/nyc_shortest_paths_3500m/", full.names = TRUE)

read_all_files <- function(pattern = pattern, path = "./"){
  map_df(list.files(path = path, pattern = pattern, full.names = TRUE), read_csv,
         col_types = list(geoid = col_character(), park_id = col_character()))
}
distances <- read_all_files(".csv", "data/nyc_shortest_paths_3500m/") %>%
  # filter out infinite paths
  filter(!is.infinite(distance)) %>%
  # add ten meters to all paths
  mutate(distance = distance + 10)
```

Now we can calculate the utility specification. The first specification uses the
distance and size terms from the Alameda County study. The second specification
adds the amenities with the utilities found by the Kinnell et al. study.

```{r logsums}
logsums <- distances %>%
  left_join(open_spaces %>% st_set_geometry(NULL), by = c("park_id" = "id")) %>%
  mutate(
    distance_miles = (distance + 10) * 0.000621371,
    u_access = -1.768 * log(distance_miles) + 0.373 * log(as.numeric(acres)),
    u_multi = u_access + 
      open_spaces$trails * 0.99 + 
      open_spaces$playgrounds * 0.26 + 
      open_spaces$courts * 0.43
  ) %>%
  group_by(geoid) %>%
  summarise(
    access_ls = log(sum(exp(u_access), na.rm = TRUE)),
    multi_ls = log(sum(exp(u_multi), na.rm = TRUE))
  ) %>%
  mutate_at(
    vars(ends_with("_ls")),
    list(~ (. - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE))
  )

bg_access <- bg_poly %>%
  left_join(logsums, by = "geoid")
```

## 10-minute buffer

We are also interested in looking at what happens when the tracts are within a 
10-minute walk of a park, the threshold established by the Trust for Public Lands.
This is interesting because whereas a limited number of tract centroids failed
to overlap with the 10-minute buffer, all of the block group centroids intersect.

## Accessibility Map
Here's a map of the values. Note that there is also a shiny app in this project
that allows the user to set different values for the logsum parameters.

```{r access_map}
# this interactive map is much easier to create and explore
pal <- colorQuantile("GnBu", c(bg_access$access_ls, bg_access$multi_ls), n = 5)
leaflet(bg_access %>% st_transform(4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(group = "access", color = ~ pal(access_ls), stroke = 1,
              label = ~as.character(round(access_ls, 2))) %>%
  addPolygons(group = "Kinner", color = ~ pal(multi_ls), stroke = 1,
              label = ~as.character(round(multi_ls, 2))) %>%
  addLayersControl(overlayGroups = c("access", "Kinner"))
```