### 2020-05-05 03_distances.Rmd
Recompute euclidean distances
This computes Euclidean distances for block groups and tract centroids to park boundaries. The file is now too large to keep in Git.

---
title: "03_distances"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
counties <- c("081", "047", "061", "005", "085")
```

A key aspect of accessibility to parks is the distance between parks and
neighborhoods. There are two different ways to calculate distance:

   - *Euclidean* distances use the straight-line distance between two points.
   - *Network* distances use the path along a highway or sidewalk network.

Both measurements have limitiations. Euclidean distances are particularly
problematic when there are rivers, interstates, or other impassible features that
substantially lengthen the distance people travel to get between points. On the
other hand, there is not lots of data suggesting that people perceive network 
travel times very well when they are choosing destinations. We originally
estimated the models for this paper with Euclidean distances, but
our reviewers suggested we examine network-based distances as well.

Another issue the reviewers identified is the spatial resolution of the analysis.
We had used tracts, because this is where the socioeconomic and health data are 
available. But block groups are more geographically precise, and it is possible
for us to compute the accessibility at that level, if we assume that the 
socioeconomics and health information is constant across the tract (this is not
a good assumption, but we can entertain it for the purposes of the response to
reviewers).

We will use the same parks dataset we assembled [previously](02_parks_data.html).

```{r open_spaces}
open_spaces <- read_rds("data/open_spaces.rds")
```

## Euclidean Distances

The geometric centroid of the tract / block group is not a good approximation of
the center point of the tract from an accessibility standpoint. Thus, we 
get the population-weighted centroid for tracts and block groups from Census.

```{r tract_centroids}
tracts_url <- "https://www2.census.gov/geo/docs/reference/cenpop2010/tract/CenPop2010_Mean_TR36.txt"

tract_centroids <- read_csv(tracts_url) %>%
  filter(COUNTYFP %in% counties) %>%
  transmute(geoid = str_c(STATEFP, COUNTYFP, TRACTCE), LATITUDE, LONGITUDE) %>%
  # convert to sf and project into the feet-based projection used in the
  # parks dataset.
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(3628) 
```

```{r bg_centroids}
bg_url <- "https://www2.census.gov/geo/docs/reference/cenpop2010/blkgrp/CenPop2010_Mean_BG36.txt"

bg_centroids <- read_csv(bg_url) %>%
  filter(COUNTYFP %in% counties) %>%
  transmute(geoid = str_c(STATEFP, COUNTYFP, TRACTCE, BLKGRPCE), LATITUDE, LONGITUDE) %>%
  # convert to sf and project into the feet-based projection used in the
  # parks dataset.
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(3628) 
```

The `st_distance()` function from the `sf` library is able to compute the 
Euclidean distance between two sets of shapes, and gets the distance between 
points and edges of polygons directly. We calculate the distance in miles, 
and constrain that the minimum distance is 0.1 miles (no one lives in a park).

```{r distance_table}
distance_table <- function(points, poly){
  dists <- st_distance(points, poly, by_element = FALSE) %>%
    units::set_units(miles) %>%
    units::drop_units()

  dists <- pmax(dists, 0.1)

  rownames(dists) <- points$geoid
  colnames(dists) <- poly$id

  as_tibble(dists, rownames = "geoid") %>%
    pivot_longer(cols = -geoid, names_to = "park_id", values_to = "distance")
}

dist_tr_euc <- distance_table(tract_centroids, open_spaces)
dist_bg_euc <- distance_table(bg_centroids,    open_spaces)
```

```{r distances}
list("tracts" = dist_tr_euc, "blockgroups" = dist_bg_euc) %>%
  write_rds("data/distances.rds")
```

## Network Distances

For this