### 2020-06-05 manuscript/submission/coverletter/coverletter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 2.0 (12/2/17)
%
% This template originates from:
% http://www.LaTeXTemplates.com
%
% Authors:
% Brian Moses
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)

\input{structure.tex} % Include the file that specifies the document structure

%\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," and date to the left of the page
\setlength\parindent{0pt}

\usepackage{fontspec}

% Classic
\setmainfont[BoldFont={Roboto Bold}]{Roboto Light}

%----------------------------------------------------------------------------------------
%	YOUR INFORMATION
%----------------------------------------------------------------------------------------

\Who{Gregory S. Macfarlane} % Your name

\Title{, PhD, PE} % Your title, leave blank for no title

\authordetails{
	%Civil \& Environmental Engineering\\ % Your department/institution
	430 Engineering Building \\ % Your address
	Provo, UT 84602\\ % Your city, zip code, country, etc
	Email: gregmacfarlane@byu.edu\\ % Your email address
	Phone: (801) 422-8505\\ % Your phone number
}

%----------------------------------------------------------------------------------------
%	HEADER CONTENTS
%----------------------------------------------------------------------------------------

\logo{logo.png} % Logo filename, your logo should have square dimensions (i.e. roughly the same width and height), if it does not, you will need to adjust spacing within the HEADER STRUCTURE block in structure.tex (read the comments carefully!)

\headerlineone{} % Top header line, leave blank if you only want the bottom line

\headerlinetwo{} % Bottom header line

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}{
}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\opening{Dear Professor See,}

We would like to thank both reviewers for their thoughtful and helpful comments. We are delighted that they found our submission a “well-written paper that makes a useful contribution to the accessibility literature regarding quantifying accessibility” as well as “a nice added-value to the extensive literature on the subject.”

We feel we have addressed each comment raised by the reviewers, both major and minor. In particular, we re-executed our analysis using network distances in place of the Euclidean distances used in the original manuscript; the findings of this analysis were inconsequential, and our decision to retain the Euclidean specification is detailed in the attached response to reviewers.

We are grateful for the opportunity to submit this revised manuscript for further review and consideration in {\em Environment and Planning B: Urban Analytics and City Science}.

\closing{Regards,}

%----------------------------------------------------------------------------------------
%	OPTIONAL FOOTNOTE
%----------------------------------------------------------------------------------------

% Uncomment the 4 lines below to print a footnote with custom text
%\def\thefootnote{}
%\def\footnoterule{\hrule}
%\footnotetext{\hspace*{\fill}{\footnotesize\em Footnote text}}
%\def\thefootnote{\arabic{footnote}}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}