### 2020-09-25 manuscript/submission/coverletter/coverletter.tex
New cover letter to editor

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 2.0 (12/2/17)
%
% This template originates from:
% http://www.LaTeXTemplates.com
%
% Authors:
% Brian Moses
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)

\input{structure.tex} % Include the file that specifies the document structure

%\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," and date to the left of the page
\setlength\parindent{0pt}

\usepackage{fontspec}

% Classic
\setmainfont[BoldFont={Roboto Bold}]{Roboto Light}

%----------------------------------------------------------------------------------------
%	YOUR INFORMATION
%----------------------------------------------------------------------------------------

\Who{Gregory S. Macfarlane} % Your name

\Title{, PhD, PE} % Your title, leave blank for no title

\authordetails{
	%Civil \& Environmental Engineering\\ % Your department/institution
	430 Engineering Building \\ % Your address
	Provo, UT 84602\\ % Your city, zip code, country, etc
	Email: gregmacfarlane@byu.edu\\ % Your email address
	Phone: (801) 422-8505\\ % Your phone number
}

%----------------------------------------------------------------------------------------
%	HEADER CONTENTS
%----------------------------------------------------------------------------------------

\logo{logo.png} % Logo filename, your logo should have square dimensions (i.e. roughly the same width and height), if it does not, you will need to adjust spacing within the HEADER STRUCTURE block in structure.tex (read the comments carefully!)

\headerlineone{} % Top header line, leave blank if you only want the bottom line

\headerlinetwo{} % Bottom header line

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}{
}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\opening{Dear Professor See,}

Please find attached the second revision of our manuscript, \emph{Modeling the
impacts of park access on health outcomes: a utility-based accessibility
approach.}  We are pleased that one of the original referees appears wholly
satisfied by  our revisions. Though we are disappointed that the other original
referee could not provide comment, we are grateful to have received additional
comments from a third referee.

Many of the comments from this new referee were extraordinarily helpful to us as
we seek to clarify the specific contribution of our work, and to ensure that it
reaches the broadest possible audience. In particular, the recommendation to remove
cemeteries from consideration modestly strengthened our econometric findings.
At the same time, we feel that the new referee made some comments that we
exhaustively addressed in our previous response to Referee 1 (using Euclidean
distances, spatial resolution), or that contradict the recommendations of the
original referee that did not re-review the manuscript (including data on park use).

An attached document provides a more detailed response to each of the new
referee's comments. We reiterate our willingness to supply more explanatory detail
on our distance and spatial resolution investigations in a supplemental file at
your invitation.

Lastly, we are grateful for the opportunity to submit this revised manuscript for
further review and consideration in {\em Environment and Planning B: Urban
Analytics and City Science}.

\closing{Regards,}

%----------------------------------------------------------------------------------------
%	OPTIONAL FOOTNOTE
%----------------------------------------------------------------------------------------

% Uncomment the 4 lines below to print a footnote with custom text
%\def\thefootnote{}
%\def\footnoterule{\hrule}
%\footnotetext{\hspace*{\fill}{\footnotesize\em Footnote text}}
%\def\thefootnote{\arabic{footnote}}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}